Open Questions for Computer Science and Cheminformatics

### 2011-11-21 OpenQuestions.tex
some clean up of open questions

\section*{Data mining}
\begin{enumerate}
\item evaluation of similarities in a heterogeneous network (JKW: What are heterogenous networks?).  What is a specific example here?
\item integrate chemical structure information with ontologies.  Specific example problem?
\item Classify molecular descriptors up to equivalence, and dependence on one another.
\item systems-level understanding of small molecules (JKW: which systems? This will help being clearer on the challenge).  What does this mean, and what would a specific challenge problem be?
%\item \emph{Efficient molecule browsing, e.g. on scaffold level}.
%
%Chemical Abstract Services have a molecule browsing tool called SubScape, which allows to browse large-scale
%chemical spaces efficiently. What could be large-scale solutions for doing this within (combined and aligned) 
%public databases.
%
\item \emph{Dynamic similarity search on instant binary vectors}.

@@

  There are various tools for doing automatic text mining on chemical
  patents. Still, the overall acceptance rate of chemical text mining
  is improvable, since many medicinal chemists are very concerned
  about the data quality of such efforts.  What could be done to
  improve the mining quality, curate the obtained data, and to provide
  confidence level estimations for each molecule coming from patent
  mining? Do require image2structure and text2structure mining also
  data stores for ensuring a sufficient amount of confidence and data
  quality?  How can patent mining be used to create new drugs faster
  or to speed-up collaboration/licensing discussions?
\end{enumerate}

\section*{Machine learning}
\begin{enumerate}
\item \emph{Large-scale vectorial versus kernels molecule similarity}

@@

\section*{Software engineering}
\begin{enumerate}
\item \emph{Large-scale exchange and use of 3D conformational databases}

  There exist a multitude of programs generation 3D conformations of
  molecules. Still, often input formats vary and require to re-create
  3D conformational spaces with other programs. This does not only
  complicate the maintenance scenario of such databases, especially
  with the explosion of public databases, but also creates hugh space
  and time complexity issues when searching within such
  databases. What could be better interfaces, maintenance, data
  structures, and private/public sharing scenarios for conformational
  3D databases?

\item \emph{Parallelization of cheminformatics problems \& algorithms}
Many cheminformatics tasks apply algorithms over large input files or
across many molecules. A simple way to parallelize this is to chunk
the input data and let individual threads/node process each chunk. A
trivial solution is to manually chunk the input and submit a series of
jobs to a scheduling system.

Can we employ modern frameworks like Hadoop to perform this? Can we
develop a generalized framework that includes the requisite
cheminformatics tools that allows users to seamlessly distribute jobs
over Hadoop and other map reduce systems?

Going further, are there cheminformatics problems that can make use of
the map/reduce paradigm at the algorithmic level? More generally, are
cheminformatics algorithms amenable to parallelization (that does not
involve chunking the input data)? Are they worth the effort?

@@

Let us assume an organization, e.g. a commercial company, has a single or a small number of established KM and ELN products. 
How can we improve the maintenance, leveraging, and collaboration with many external partners (each of them potentially
with another KM/ELN solution)? Which party is hosting which data in which data structure (ontologies?), and 
how can we ensure that only pre-defined data entries 
(and a limited number of annotations, e.g. biological activities) are visible to a partner.
How can this be organized for a multitude of partners? Cloud computing, user management, encryption granularity and efficient security management?
\end{enumerate}

>>>>

\section*{Data mining \& machine learning}
\begin{enumerate}
\item evaluation of similarities in a heterogeneous network (JKW: What are heterogenous networks?).  What is a specific example here?
\item integrate chemical structure information with ontologies.  Specific example problem?
\item Classify molecular descriptors up to equivalence, and dependence on one another.
\item systems-level understanding of small molecules (JKW: which systems? This will help being clearer on the challenge).  What does this mean, and what would a specific challenge problem be?
%\item \emph{Efficient molecule browsing, e.g. on scaffold level}.
%
%Chemical Abstract Services have a molecule browsing tool called SubScape, which allows to browse large-scale
%chemical spaces efficiently. What could be large-scale solutions for doing this within (combined and aligned) 
%public databases.
%
\item \emph{Dynamic similarity search on instant binary vectors}.

@@

  There are various tools for doing automatic text mining on chemical
  patents. Still, the overall acceptance rate of chemical text mining
  is improvable, since many medicinal chemists are very concerned
  about the data quality of such efforts.  What could be done to
  improve the mining quality, curate the obtained data, and to provide
  confidence level estimations for each molecule coming from patent
  mining? Do require image2structure and text2structure mining also
  data stores for ensuring a sufficient amount of confidence and data
  quality?  How can patent mining be used to create new drugs faster
  or to speed-up collaboration/licensing discussions?

\item \emph{Large-scale vectorial versus kernels molecule similarity}

@@

\section*{Databases \& software engineering}
\begin{enumerate}
\item \emph{Large scale conformational databases}

  While many programs are availabel to generate 3D conformations, on
  the fly generation for large collections can be time
  consuming. Rather, can we store massive conformer collections and
  support 3D searches over them? Beyond generating the conformers
  themselves, this problem has several aspects covering database
  design, parallel systems and software engineering:
  \begin{itemize}
  \item How many conformers are required to provide \emph{sufficient} coverage?
  \item What representation will be used to store conformers and run
    queries? How does the choice of representation affect the type of
    queries we can run?
  \item Most 3D similarity approaches either employ a vectorial
    representation or a volumetric representation. How can these be
    efficiently indexed?
  \item What type of parallel infrastructure can be used to speed up queries?
  \end{itemize}

\item \emph{Map/Reduce in cheminformatics}
  Many cheminformatics tasks apply algorithms over large input files
  or across many molecules (similarity and substructure search,
  docking, filtering). A simple way to parallelize this is to chunk
  the input data and let individual threads/nodes process each chunk. A
  trivial solution is to manually chunk the input and submit a series
  of jobs to a scheduling system. Using the Hadoop ecosystem as an
  example, we can ask various questions:
  \begin{itemize}
  \item Can we employ modern frameworks like Hadoop to support
    embarassingly parallel cases (filtering $10^7$ molecule libraries)?
  \item Can we develop a generalized framework that includes the
    requisite cheminformatics tools that allows users to seamlessly
    distribute jobs over Hadoop and other map reduce systems?
  \item Is the latency involved in Hadoop based datastores (HBase)
    worth the ability to handle massive molecule collections and run
    M/R queries across them? 
  \item Going further, are there cheminformatics problems that can make use
  of the map/reduce paradigm at the algorithmic level?
  \end{itemize}
\end{enumerate}

  Let us assume an organization, e.g. a commercial company, has a
  single or a small number of established KM and ELN products.  How
  can we improve the maintenance, leveraging, and collaboration with
  many external partners (each of them potentially with another KM/ELN
  solution)? Which party is hosting which data in which data structure
  (ontologies?), and how can we ensure that only pre-defined data
  entries (and a limited number of annotations, e.g. biological
  activities) are visible to a partner.  How can this be organized for
  a multitude of partners? Cloud computing, user management,
  encryption granularity and efficient security management?

@@
